---
title: "Spatial cueing effects in a go/no-go task"
author: "Pierce Edmiston"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial cueing effects in a go/no-go task}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, config, echo = FALSE}
library(gridExtra)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

devtools::load_all()
```

```{r, get-spatial-cueing-data}
cueing <- go_nogo
```

# Methods

## Participants

```{r, num-subjs-in-each-experiment}
cueing %>% group_by(subj_id, interval, flicker) %>%
  summarize(num_trials = n()) %>%
  count(interval, flicker)
```

## Procedure

## Analysis

Fit a model to predict accuracy following each cue type in each experiment condition.
`flicker` and `interval` were varied between-subjects. `cue type` was
varied within-subjects. Trials were evenly distributed between frame cues,
sound cues, and no cue baseline trials. To estimate the effect of each
cue (frame, sound) relative to baseline performance, two contrast variables
were created. These contrast variables (`frame_v_nocue` and `sound_v_nocue`)
are interpreted as the independent cueing effects of frames and sounds.

We are testing the hypothesis that the cueing effects are differentially
affected by the flicker and interval manipulations. In particular, we
expect the `frame_v_nocue` cueing effect to be more affected by the
flicker manipulation than the `sound_v_nocue` cueing effect.

The contention is that it is only in a relatively small window of conditions
that visual cues "outperform" so-called endogenous cues. For a majority
of the visual tasks in our everyday lives, we are constantly confronted
by irrelevant information and distractions. When there are distractions,
such as a flickering visual mask, exogenous cues are not as effective for
achieving heightened visual awareness.

# Results

## Cueing effects on accuracy

```{r, fit-accuracy-mod}
library(lme4)

contrast_cols <- c("frame_v_nocue", "sound_v_nocue")
cueing[,contrast_cols] <- create_cue_type_treatment_contrasts(cueing$cue_type)
cueing$flicker_c <- center_flicker(cueing$flicker)
cueing$interval_c <- center_interval(cueing$interval)

accuracy_mod <- glmer(is_correct ~ (frame_v_nocue + sound_v_nocue) * (flicker_c + interval_c) +
                        (frame_v_nocue + sound_v_nocue || subj_id),
                      family = "binomial", data = cueing)
summary(accuracy_mod)
```

(discussion of model results)

```{r, create-between-subj-conditions-table}
between_subj_conditions <- expand.grid(
    flicker = c("on", "off"),
    interval = c(0.75, 0.10)
  )
between_subj_conditions$exp <- row.names(between_subj_conditions)
between_subj_conditions
```

```{r, estimate-cueing-effect-in-accuracies-in-each-condition}
library(broom)
library(purrr)

cueing_effects_in_accuracies <- between_subj_conditions %>%
  # for each experiment
  split(.$exp) %>%

  # fit a model to predict the cueing effects
  map(~ glmer(is_correct ~ frame_v_nocue + sound_v_nocue + (frame_v_nocue + sound_v_nocue || subj_id),
              data = filter(cueing, flicker == .$flicker, interval == .$interval),
              family = "binomial")
  ) %>%

  # extract the parameter estimates from the model
  map(tidy, effects = "fixed") %>%
  map(filter, term %in% c("frame_v_nocue", "sound_v_nocue")) %>%

  # relabel the between subject conditions
  rbind_all(.) %>%
  mutate(
    flicker = rep(between_subj_conditions$flicker, each = 2),
    interval = rep(between_subj_conditions$interval, each = 2)
  )

cueing_effects_in_accuracies
```

```{r, cueing-effects-in-accuracies}
library(ggplot2)
library(scales)

cueing_effects_in_accuracies$line_group <- create_line_groups(cueing_effects_in_accuracies)
cueing_effects_in_accuracies$flicker_label <- label_flicker(cueing_effects_in_accuracies$flicker)

cueing_effects_in_accuracies_plot <- ggplot(cueing_effects_in_accuracies,
    aes(x = interval, y = estimate)) +
  geom_pointrange(aes(ymin = estimate - std.error, ymax = estimate + std.error, color = term)) +
  geom_line(aes(group = line_group, color = term, lty = flicker)) +
  facet_wrap("flicker_label", nrow = 2) +
  geom_hline(yintercept = 0.0, lty = 2) +
  scale_interval_axis +
  scale_y_continuous("Cueing Effect in Accuracy (log-odds)") +
  scale_term_legend +
  coord_cartesian(xlim = c(0.0, 1.0)) +
  guides(lty = "none") +
  ggtitle("Cueing Effects in Accuracies")
cueing_effects_in_accuracies_plot
```

## Cueing effects on reaction times

```{r, fit-rt-mod}
hits <- filter(cueing, target_present == 1, response == "go")

rt_mod <- lmer(rt ~ (frame_v_nocue + sound_v_nocue) * (flicker_c + interval_c) +
                 (frame_v_nocue + sound_v_nocue || subj_id), data = hits)
summary(rt_mod)
```

(dicussion of model results)

```{r, estimate-cueing-effect-in-rts-in-each-condition}
cueing_effects_in_rts <- between_subj_conditions %>%
  # for each experiment
  split(.$exp) %>%

  # fit a model to predict the cueing effects
  map(~ lmer(rt ~ frame_v_nocue + sound_v_nocue + (frame_v_nocue + sound_v_nocue || subj_id),
              data = filter(hits, flicker == .$flicker, interval == .$interval))
  ) %>%

  # extract the parameter estimates from the model
  map(tidy, effects = "fixed") %>%
  map(filter, term %in% c("frame_v_nocue", "sound_v_nocue")) %>%

  # relabel the between subject conditions
  rbind_all(.) %>%
  mutate(
    flicker = rep(between_subj_conditions$flicker, each = 2),
    interval = rep(between_subj_conditions$interval, each = 2)
  )

cueing_effects_in_rts
```

```{r, cueing-effect-in-rts}
cueing_effects_in_rts$line_group <- create_line_groups(cueing_effects_in_rts)
cueing_effects_in_rts$flicker_label <- label_flicker(cueing_effects_in_rts$flicker)

cueing_effects_in_rts_plot <- ggplot(cueing_effects_in_rts, aes(x = interval, y = estimate)) +
  geom_pointrange(aes(ymin = estimate - std.error, ymax = estimate + std.error, color = term)) +
  geom_line(aes(group = line_group, color = term, lty = flicker)) +
  facet_wrap("flicker_label", nrow = 2) +
  geom_hline(yintercept = 0.0, lty = 2) +
  scale_interval_axis +
  scale_y_continuous("Cueing Effect in RTs (ms)") +
  scale_term_legend +
  coord_cartesian(xlim = c(0.0, 1.0)) +
  guides(lty = "none") +
  ggtitle("Cueing Effects in RTs")
cueing_effects_in_rts_plot
```

```{r, cueing-effects-with-mask, fig.width=5, fig.height=4}
cueing_with_flicker <- cueing %>% filter(flicker == "on")
ggplot(cueing_with_flicker, aes(x = interval, y = is_correct, color = cue_type)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  scale_y_continuous("Accuracy", labels = percent) +
  theme_classic()

ggplot(cueing_with_flicker, aes(x = interval, y = rt, color = cue_type)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  scale_y_continuous("RT") +
  theme_classic()
```

```{r, cueing-effects-without-mask, fig.width=5, fig.height=4}
cueing_without_flicker <- cueing %>% filter(flicker == "off")
ggplot(cueing_without_flicker, aes(x = interval, y = is_correct, color = cue_type)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  scale_y_continuous("Accuracy", labels = percent) +
  theme_classic()

ggplot(cueing_without_flicker, aes(x = interval, y = rt, color = cue_type)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  scale_y_continuous("RT") +
  theme_classic()
```

# Discussion

```{r, means-plot, fig.height = 10, fig.width = 12}
cueing$flicker_label <- label_flicker(cueing$flicker)
rt_plot <- ggplot(cueing, aes(x = interval, y = rt, color = cue_type)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  facet_wrap("flicker_label", nrow = 2)

acc_plot <- ggplot(cueing, aes(x = interval, y = is_correct, color = cue_type)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") +
  facet_wrap("flicker_label", nrow = 2)

grid.arrange(acc_plot + theme(legend.position = "none"),
             rt_plot + theme(legend.position = c(0.92, 0.3)),
             ncol = 2)
```

```{r, results-plot, fig.height = 10, fig.width = 12}
grid.arrange(cueing_effects_in_accuracies_plot + theme(legend.position = "none"),
             cueing_effects_in_rts_plot + theme(legend.position = c(0.92, 0.8)),
             ncol = 2)
```

The experiments reported here do not establish why exogenous and endogenous
cues differ in this way, a question to be addressed in future work. They
do, however, motivate that work.

Traditionally, the difference between exogenous and endogenous cues has been
attributed to time. Exogenous cues are faster acting than
endogenous cues. There is also an element of volition: exogenous cues are
automatic, whereas endogenous cues can be discounted.

But an implicit assumption in the literature is that controlling for these
other factors, exogenous and endogenous cues can mean roughly the same thing
when it comes to cueing a particular spatial location. Under the hood it is
the same "spotlight of attention" in both cueing conditions.

Of course, we know more about what neural states result in
attended versus unattended behaviors, and it
